---
title: 'Bios 6301: Assignment 5'
output: pdf_document
---

*Due Thursday, 12 October, 1:00 PM*

$5^{n=day}$ points taken off for each day late.

40 points total.

Submit a single knitr file (named `homework5.rmd`), along with a valid PDF output file. Inside the file, clearly indicate which parts of your responses go with which problems (you may use the original homework document as a template). Add your name as `author` to the file's metadata section. Raw R code/output or word processor files are not acceptable.

Failure to name file `homework5.rmd` or include author name may result in 5 points taken off.

### Question 1 ###

**15 points**

A problem with the Newton-Raphson algorithm is that it needs the derivative $f'$. If the derivative is hard to compute or does not exist, then we can use the *secant method*, which only requires that the function $f$ is continuous.

Like the Newton-Raphson method, the **secant method** is based on a linear approximation to the function $f$. Suppose that $f$ has a root at $a$. For this method we assume that we have *two* current guesses, $x_0$ and $x_1$, for the value of $a$. We will think of $x_0$ as an older guess and we want to replace the pair $x_0$, $x_1$ by the pair $x_1$, $x_2$, where $x_2$ is a new guess.

To find a good new guess x2 we first draw the straight line from $(x_0,f(x_0))$ to $(x_1,f(x_1))$, which is called a secant of the curve $y = f(x)$. Like the tangent, the secant is a linear approximation of the behavior of $y = f(x)$, in the region of the points $x_0$ and $x_1$. As the new guess we will use the x-coordinate $x_2$ of the point at which the secant crosses the x-axis.

The general form of the recurrence equation for the secant method is:

$$x_{i+1} = x_i - f(x_i)\frac{x_i - x_{i-1}}{f(x_i) - f(x_{i-1})}$$

Notice that we no longer need to know $f'$ but in return we have to provide *two* initial points, $x_0$ and $x_1$.

**Write a function that implements the secant algorithm.** Validate your program by finding the root of the function $f(x) = \cos(x) - x$. Compare its performance with the Newton-Raphson method -- which is faster, and by how much?  For this example $f'(x) = -\sin(x) - 1$.


#Secant Method
```{r}
x0 <- 1
x1 <- 0.5
f<- function(x) {cos(x) - x}  
plot(f, xlim = c(0.5, 1))
secant_method<- function(f, x0, X1, tol=1e-9, max_iter=1000){ 
  iter<- 0
    while((abs(x1-x0)) >tol) {
    x2<- x1-(f(x1)*(x1-x0))/ (f(x1)-f(x0))
    x0<-x1
    x1<-x2 
    iter<- iter+1
    if (iter>max_iter) break 
    }
     return (c(x2, iter)) 
   }
secant_method(f,x0,x1)
```


```{r}
#Newton raphson algorithm
start.time <- Sys.time()
f<- function(x) {cos(x) - x}
d<- function(x) {-sin(x) - 1}
x <- 0.5
newton_raphson <-function(f, x, d, tol=1e-9, n=1000) { 
  i <- 0
  while(abs(f(x)) > tol){
     x<- x - (f(x)/d(x)) 
    i<- i+1 
    if (i>n) break 
  }
  x
}
newton_raphson(f,x,d)
end.time <- Sys.time()
time.taken <- end.time - start.time
time.taken  
```

### Question 2 ###

**20 points**

The game of craps is played as follows (this is simplified). First, you roll two six-sided dice; let x be the sum of the dice on the first roll. If x = 7 or 11 you win, otherwise you keep rolling until either you get x again, in which case you also win, or until you get a 7 or 11, in which case you lose.

Write a program to simulate a game of craps. You can use the following snippet of code to simulate the roll of two (fair) dice:

```{r}
    x <- sum(ceiling(6*runif(2)))
```

1. The instructor should be able to easily import and run your program (function), and obtain output that clearly shows how the game progressed.  Set the RNG seed with `set.seed(100)` and show the output of three games. (lucky 13 points)
```{r}
set.seed(100)
fcraps = function() {
  x = sum(ceiling(6*runif(2)))
  if (x==7|x==11) return("win")
  while (TRUE) {
    x1 = sum(ceiling(6*runif(2)))
    if (x1 == x) return("win")
    if (x1 == 7|x1==11) return("lose")
  }
}
ncraps = function(f=fcraps,n){
  res = NULL
  for (i in 1:n){
    res[i]=f()
  }
  return(res)
}
ncraps(n=3)
```
1. Find a seed that will win ten straight games.  Consider adding an argument to your function that disables output. Show the output of the ten games. (7 points)
```{r}
seed = 1

while (seed<10000){
  set.seed(seed)
  res = ncraps(n=10)
  a = rep('win',10)
  if (length(unique(res))==1){
    if ((res %in% 'win')[1]==TRUE) break
  }
  seed=seed+1
}
seed
set.seed(seed)
ncraps(n=10)
```
### Question 3 ###

**5 points**

This code makes a list of all functions in the base package:

```{r eval=FALSE}
objs <- mget(ls("package:base"), inherits = TRUE)
funs <- Filter(is.function, objs)
```

Using this list, write code to answer these questions.

1. Which function has the most arguments? (3 points)
```{r}
# find number of arguments for each functions
objs <- mget(ls("package:base"), inherits = TRUE)
funs <- Filter(is.function, objs)
args.number <- sapply(funs, function(x){length(formalArgs(x))})
max(args.number)
# find the function with the most arguments
funs[which.max(args.number)]
```
1. How many functions have no arguments? (2 points)
```{r}
no.args <- (args.number == 0)
sum(no.args)
```
Hint: find a function that returns the arguments for a given function.
